---
title: "Demo for \"Core-Elements for Classical Linear Regression\""
date: "2023-05-18"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, include = FALSE}
rm(list = ls())
gc()
```

This is a demo for implementing the proposed core-elements method and comparing it with other competitors, under the case of **D1, R1** without data corruption.

## Preparation
First, we import some R Packages and a cpp source file. The source file is used for sparse matrix operations.
```{r}
list.of.packages <- c("mvtnorm", "MASS", "corpcor", "pracma", "Rfast", "Rcpp", "Matrix", "ggplot2")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
if (length(new.packages)) install.packages(new.packages)
library(mvtnorm)
library(MASS)
library(corpcor)
library(pracma)
library(Rfast)
library(Rcpp)
library(Matrix)
library(ggplot2)

setwd(getwd())
sourceCpp("source/funcs.cpp")
```

Next, we define some functions:

* `yGenerate`: generate the response vector $\boldsymbol{y}$;

* `xSparse`: sparsify the predictor matrix $\mathbf{X}$.
```{r}
yGenerate <- function(x, beta, SNR) {
  # generate the response vector
  y_raw <- as.vector(x %*% beta)
  N <- length(y_raw)
  y <- y_raw + rnorm(N, 0, sd(y_raw)/sqrt(SNR))
  return(y)
}

xSparse <- function(x, sparsity) {
  # sparsify the predictor matrix
  N <- length(x[, 1])
  p <- length(x[1, ])
  xx <- rep(0, N * p)
  x_vec <- as.vector(x)
  id <- sample(1:(N * p), (1 - sparsity) * N * p, replace = FALSE)  # non-zeros
  xx[id] <- x_vec[id]
  xx <- matrix(xx, N, p)
  return(xx)
}
```

Then, we set parameters for this experiment.
```{r}
set.seed(1000)
case_type <- "D1"  # normal distribution
spar_type <- "R1"  # dense design matrix
case <- c(case_type, spar_type)
N <- 10000  # sample size n
p <- 100  # dimension p
sub_meta <- (1:5) * 2 * p  # subsample parameter r
nloop <- 20  # number of replicates
SNR <- 4  # signal-to-noise ratio
num_method <- 5
beta_true <- rep(1, p)  # true coefficients
```

## Estimation
Now, we use different approaches to estimate the coefficient vector $\boldsymbol{\beta}$. The competitors include UNIF, BLEV, SLEV, IBOSS, and the proposed CORE method.

```{r}
mse_meta <- array(0, dim = c(nloop, num_method, length(sub_meta)))
mse_temp <- matrix(0, num_method, length(sub_meta))

## Beginning ##
for (i in 1:nloop) {
  
  ## Set the covariance matrix of X ##
  Sigma <- diag(p)
  for (ii in 1:p) {
    for (jj in 1:p) {
      Sigma[ii, jj] <- 0.6^(abs(ii - jj))
    }
  }
  set.seed(100 + 123 * i)
  
  ## Generate the design matrix X ##
  if (case[1] == "D1") {
    X <- mvrnorm(N, rep(0, p), Sigma)
  } else if (case[1] == "D2") {
    X <- mvrnorm(N, rep(0, p), Sigma)
    X <- exp(X)
    X <- X - matrix(rep(colMeans(X), N), N, p, byrow = TRUE)
  } else if (case[1] == "D3") {
    X <- mvtnorm::rmvt(N, Sigma, delta = rep(0, p), df = 3)
  }
  
  ## Sparsify the design matrix X ##
  if (case[2] == "R2") {
    X <- xSparse(X, 0.2)
  } else if (case[2] == "R3") {
    X <- xSparse(X, 0.4)
  } else if (case[2] == "R4") {
    X <- xSparse(X, 0.6)
  } else if (case[2] == "R5") {
    X <- xSparse(X, 0.8)
  }
  
  id_nz <- which(rowSums(abs(X)) != 0)
  X <- X[id_nz, ]
  Y <- yGenerate(X, beta_true, SNR)  # generate the response y
  N <- length(Y)
  
  ## Compute the sampling probabilities for BLEV and SLEV ##
  U <- fast.svd(X)$u
  PP <- rowSums(U^2)
  prob_blev <- PP/p
  prob_slev <- 0.9 * prob_blev + 0.1/N
  rm(PP, U)
  
  for (j in 1:length(sub_meta)) {
    set.seed(100 + 421 * j + 123 * i)
    sub <- sub_meta[j]
    
    ## Subsampling methods ##
    # UNIF
    id <- sample(N, sub, replace = FALSE)
    xt <- X[id, ]
    yt <- Y[id]
    fit <- lm(yt ~ xt - 1)
    unif_temp <- fit$coefficients
    rm(id, xt, yt, fit)
    
    # BLEV
    id <- sample(N, sub, replace = FALSE, prob_blev)
    xt <- X[id, ]
    yt <- Y[id]
    wgt <- 1/prob_blev[id]
    fit <- lm(yt ~ xt - 1, weights = wgt)
    blev_temp <- fit$coefficients
    rm(id, xt, yt, wgt, fit)
    
    # SLEV
    id <- sample(N, sub, replace = FALSE, prob_slev)
    xt <- X[id, ]
    yt <- Y[id]
    wgt <- 1/prob_slev[id]
    fit <- lm(yt ~ xt - 1, weights = wgt)
    slev_temp <- fit$coefficients
    rm(id, xt, yt, wgt, fit)
    
    # IBOSS
    id <- NULL
    X_c <- X
    ssub <- floor(sub/p/2)
    
    for (k in 1:p) {
      X_c[id, ] <- NA
      thres1 <- Rfast::nth(X_c[, k], ssub, descending = TRUE)
      id1 <- which(X_c[, k] >= thres1)[1:ssub]
      thres2 <- Rfast::nth(X_c[, k], ssub, descending = FALSE)
      id2 <- which(X_c[, k] <= thres2)[1:ssub]
      id <- c(id, id1, id2)
    }
    xt <- X[id, ]
    yt <- Y[id]
    
    fit <- lm(yt ~ xt - 1)
    iboss_temp <- fit$coefficients
    rm(X_c, id1, id2, id, xt, yt, fit)
    
    # CORE
    XX <- matrix(0, N, p)
    idd <- NULL
    
    thres <- colnth(abs(X), rep(sub, p), descending = TRUE, parallel = FALSE)
    for (kk in 1:p) {
      col <- abs(X[, kk])
      id <- which(col >= thres[kk])[1:sub]
      XX[id, kk] <- X[id, kk]
      idd <- unique(c(idd, id))
    }
    xxt <- XX[idd, ]
    yyt <- Y[idd]
    xt <- X[idd, ]
    
    xt1 <- Matrix(xt, sparse = TRUE)
    xxt1 <- Matrix(xxt, sparse = TRUE)
    core_temp <- eigenMultSolveSp(xt1, xxt1, yyt)
    rm(XX, col, thres, id, idd, xt, xxt, yyt, xt1, xxt1)
    
    
    ######################################################
    coef <- cbind(unif_temp, blev_temp, slev_temp, iboss_temp, core_temp)
    coef[is.na(coef)] <- 0
    coef_mse <- colMeans((coef - beta_true)^2)/mean(beta_true^2)
    mse_temp[, j] <- coef_mse
  }
  
  mse_meta[i, , ] <- mse_temp
}

UNIF <- log(mse_meta[, 1, ])
BLEV <- log(mse_meta[, 2, ])
SLEV <- log(mse_meta[, 3, ])
IBOSS <- log(mse_meta[, 4, ])
CORE <- log(mse_meta[, 5, ])
```

## Visualization

At last, we plot the MSE of different methods for comparison.
```{r}
mse_mat <- data.frame(mse = c(apply(UNIF, 2, mean), apply(BLEV, 2, mean), apply(SLEV, 2, mean), 
                              apply(IBOSS, 2, mean), apply(CORE, 2, mean)), 
                      sd = c(apply(UNIF, 2, sd), apply(BLEV, 2, sd), apply(SLEV, 2, sd), 
                             apply(IBOSS, 2, sd), apply(CORE, 2, sd)), 
                      Method = factor(rep(c("UNIF", "BLEV", "SLEV", "IBOSS", "CORE"), 
                                          each = length(sub_meta))), 
                      sub = rep(sub_meta/p, num_method))
mse_mat$Method <- factor(mse_mat$Method, levels = c("UNIF", "BLEV", "SLEV", "IBOSS", "CORE"))

pd <- position_dodge(0.4)
p1 <- ggplot(mse_mat, aes(x = sub, y = mse, group = Method, colour = Method))
p23 <- p1 + theme_bw() + theme(panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(), 
                               panel.border = element_rect(colour = "black"),
                               axis.text = element_text(size = 19), 
                               axis.title = element_text(size = 19),
                               legend.position = "right", 
                               legend.title = element_blank(),
                               legend.background = element_rect(fill = alpha("white", 0.6), color = "black"), 
                               legend.key.width = unit(3, "line"), 
                               legend.key.height = unit(1, "line"), 
                               legend.text = element_text(size = 13)) + 
  geom_errorbar(aes(ymin = mse - sd, ymax = mse + sd), width = 1, position = pd, show.legend = FALSE) +
  geom_line(aes(linetype = Method, size = Method), position = pd) +
  geom_point(position = pd, aes(shape = Method), size = 2) +
  scale_shape_manual(values = c(8, 4, 5, 7, 1)) + 
  scale_linetype_manual(values = c(5, 4, 3, 2, 1)) + 
  scale_size_manual(values = c(1, 1, 1, 1, 1)) + 
  scale_color_manual(values = c("#999999", "#0033CC", "#3399FF", "orchid", "red")) + 
  labs(title = paste(case, collapse = ", "), x = "r/p", y = "log(MSE)") + 
  theme(plot.title = element_text(hjust = 0.5, size = 19))
print(p23)
```
